---
author: "Matthieu VERON"
date: "6/11/2022"
output:
    pdf_document:
    fig_caption: yes
    fig_height: 4
    fig_width: 6
header-includes:   
-  \pagenumbering{gobble}
number_sections: yes
csl: fish-and-fisheries.csl
toc: yes
title: "A template to compare the latest version of GMACS with the last available assesment."
---



```{r setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(echo=FALSE,message=FALSE,warning=FALSE)
main.path <- dirname(dirname(rstudioapi::getActiveDocumentContext()$path))

knitr::opts_chunk$set(root.dir = main.path)
knitr::opts_knit$set(root.dir = main.path)
```




```{r}

rm(list = ls())

# setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

library(gmr)
```

\newpage

```{r Define runs and versions, echo=F, message=FALSE, warning=FALSE, include=FALSE}

# Species of interest
.Spc <-c(
  # "all"
  # "EAG",
  # "WAG",
  # "BBRKC",
  "SMBKC"
  # "SNOW_crab"
  )

# Names of the GMACS version to consider for run
.GMACS_version <- c(
  "Last_Assessment",
  "Dvpt_Version"
  )

# Define directory
.VERSIONDIR <- c(
  paste0(getwd(), "/Assessments/"),
  paste0(getwd(), "/GMACS_versions/Dvpt_Version/")
)

# Use Last Assessment for comparison?
# If yes, you must provide the names of the model for each species in the variable .ASSMOD_NAMES
# Those model folder must have to be hold in the folder Assessments
.ASS <- TRUE

# names of the model for the last assessment - Only useful if comparison is made.
# if all stocks are considered they have to be ordered as follow:
# "AIGKC/EAG" / "AIGKC/WAG" / "BBRKC" / "SMBKC" / "SNOW"
.ASSMOD_NAMES <- c(
                  # "model_21_1e",
                   # "model_21_1e",
                   # "model_21_1",
                   "model_16_0"
                   # "model_21_g"
                   )

# Need to conpile the model?
  # vector of length(.GMACS_version)
  # 0: GMACS is not compiled. This assumes that an executable exists in the directory of the concerned version.
  # 1: GMACS is compiles
.COMPILE <- c(0,0)

# Run GMACS
.RUN_GMACS <- FALSE

# Use latets available data for the assessment?
.LastAssDat <- TRUE

# Define the directories for ADMB
.ADMBpaths <- "ADpaths.txt"

# Do comparison?
.MAKE_Comp <- TRUE

# Show Rterminal
.VERBOSE <- FALSE

```


\newpage


This document presents a comparison between the latest available assessment results and the latest available version of GMACS for all stocks currently assessed (or intended to be assessed) with the _**Generalized Assessment Model for Crustaceans**_.

This comparison covers various quantities and includes graphs where useful. This includes:

1. Management quantities for each version

2. Estimated recruitment time series

3. Estimated recruitment length distribution

4. Model fit to the observed mature biomass

5. Model fits to catch data (catch and bycatch)

6. Model fits to size composition data (in catch and surveys)

7. Model predicted fishing mortalities and selectivities for all sources of mortality

8. Abundances of immature and mature crab in the last year


The latest available version of GMACS is version 2.01.K - It was last compiled on 06/06/2022.
Here is a summary of the latest changes from the version 2.01.J:

  * Added an option to select how to calculate the average recruitment used in the computation of the reference points.
  This involves to options:
    1. Use an average recruitment over a given period of time, or
    2. Consider the estimated average recruitment estimate for the current year.



```{r, echo=FALSE, results='asis'}

tables <- GMACS(
  Spc = .Spc,
  GMACS_version = .GMACS_version,
  Dir = .VERSIONDIR,
  ASS = .ASS,
  AssMod_names = .ASSMOD_NAMES,
  compile = .COMPILE,
  run = .RUN_GMACS,
  LastAssDat = .LastAssDat,
  ADMBpaths = .ADMBpaths,
  make.comp = .MAKE_Comp,
  verbose <- .VERBOSE
)
# print(tables)



```



\newpage

# A.  Aleutian Islands Golden King Crab (AIGKC; *Lithodes aequispinus*) Stock Assessments
## west of 174° W longitude (WAG)
### Management quantities for each version
### Estimated recruitment time series
### Estimated recruitment length distribution
### Model fit to the observed mature biomass
### Model fits to catch data (catch and bycatch)
### Model fits to size composition data (in catch and surveys)
### Model predicted fishing mortalities and selectivities for all sources of mortality
### Abundances of immature and mature crab in the last year

## East of 174° W longitude (EAG)
### Management quantities for each version
### Estimated recruitment time series
### Estimated recruitment length distribution
### Model fit to the observed mature biomass
### Model fits to catch data (catch and bycatch)
### Model fits to size composition data (in catch and surveys)
### Model predicted fishing mortalities and selectivities for all sources of mortality
### Abundances of immature and mature crab in the last year

\newpage


# B.  Saint Matthew Island Blue King Crab (SMBKC; *Paralithodes platypus*) Stock Assessment
### Management quantities for each version
### Estimated recruitment time series
### Estimated recruitment length distribution
### Model fit to the observed mature biomass
### Model fits to catch data (catch and bycatch)
### Model fits to size composition data (in catch and surveys)
### Model predicted fishing mortalities and selectivities for all sources of mortality
### Abundances of immature and mature crab in the last year

\newpage


# C.  Bristol Bay Red King Crab (BBRKC; *Paralithodes camtschaticus*) Stock Assessment
### Management quantities for each version
### Estimated recruitment time series
### Estimated recruitment length distribution
### Model fit to the observed mature biomass
### Model fits to catch data (catch and bycatch)
### Model fits to size composition data (in catch and surveys)
### Model predicted fishing mortalities and selectivities for all sources of mortality
### Abundances of immature and mature crab in the last year


\newpage

# D.  Snow crab (*Chionoecetes opilio*) Stock Assessment
### Management quantities for each version
### Estimated recruitment time series
### Estimated recruitment length distribution
### Model fit to the observed mature biomass
### Model fits to catch data (catch and bycatch)
### Model fits to size composition data (in catch and surveys)
### Model predicted fishing mortalities and selectivities for all sources of mortality
### Abundances of immature and mature crab in the last year


\newpage

# E.  Tanner Crab (*Chionoecetes bairdi*)
### Management quantities for each version
### Estimated recruitment time series
### Estimated recruitment length distribution
### Model fit to the observed mature biomass
### Model fits to catch data (catch and bycatch)
### Model fits to size composition data (in catch and surveys)
### Model predicted fishing mortalities and selectivities for all sources of mortality
### Abundances of immature and mature crab in the last year


\newpage

# F.  Norton Sound Red King Crab Stock Assessment (*Paralithodes camtschaticus*)
### Management quantities for each version
### Estimated recruitment time series
### Estimated recruitment length distribution
### Model fit to the observed mature biomass
### Model fits to catch data (catch and bycatch)
### Model fits to size composition data (in catch and surveys)
### Model predicted fishing mortalities and selectivities for all sources of mortality
### Abundances of immature and mature crab in the last year






































