---
title: "A basic workflow for submtting a new version of GMACS n GitHub"
author: "Matthieu VERON"
date: "1/1/2023"
output:
  pdf_document:
    number_sections: yes
  html_notebook:
    number_sections: yes
  word_document:
    number_section: yes
  html_document:
    df_print: paged
---

```{r, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE,
  include = TRUE
)

main.path <- dirname(rstudioapi::getActiveDocumentContext()$path)
knitr::opts_chunk$set(root.dir = main.path)
knitr::opts_knit$set(root.dir = main.path)

```



[Github Desktop]: https://desktop.github.com/
["Happy Git and Github for the useR"]: https://happygitwithr.com/
[R Workflow]: https://rverse-tutorials.github.io/RWorkflow-NWFSC-2021/index.html
[Github]: https://github.com/
[R]: https://cran.r-project.org/
[Rstudio]: https://www.rstudio.com/products/rstudio/download/
[`devtools`]: https://www.r-project.org/nosvn/pandoc/devtools.html
[ADMB]: https://www.admb-project.org/

[`gmr`]: https://github.com/GMACS-project/gmr
[UpdateGMACS.R]: https://github.com/GMACS-project/GMACS_Assessment_code/blob/main/GMACS/UpdateGMACS.R
[GMACS-project]: https://github.com/GMACS-project
[GMACS_Assessment_code]: https://github.com/GMACS-project/GMACS_Assessment_code
[organization]: https://github.com/GMACS-project
[GMACS]: https://github.com/GMACS-project/GMACS_Assessment_code/tree/main/GMACS
[Dvpt_Version]: https://github.com/GMACS-project/GMACS_Assessment_code/tree/main/GMACS/Dvpt_Version
[Latest_Version]: https://github.com/GMACS-project/GMACS_Assessment_code/tree/main/GMACS/Latest_Version
[ADpaths.txt]: https://github.com/GMACS-project/GMACS_Assessment_code/blob/main/GMACS/ADpaths.txt
[Assessments]: https://github.com/GMACS-project/GMACS_Assessment_code/tree/main/Assessments
[Run_GMACS]: https://github.com/GMACS-project/GMACS_Assessment_code/blob/main/GMACS/Run_GMACS.Rmd
[Compare_Version_VS_Last_Assessment]: https://github.com/GMACS-project/GMACS_Assessment_code/blob/main/GMACS/Compare_Version_VS_Last_Assessment.Rmd





# Introduction

This document is intended to give you "guidelines and generic steps" to follow when working on GMACS and releasing a new version. Its aims are to show you how to use the GMACS Github repository to modify GMACS while tracking code changes. The idea is to make this workflow easy to follow so we will work with Github without the command-line interface using the graphical interface for Git *[Github Desktop]*.

We assume you already know how to use Rstudio with projects. If you have never used Github (and/or Githhub Desktop) with Rstudio, you can find out more from the following online workshops: the ["Happy Git and Github for the useR"] from *Jenny Brian* and the [R Workflow] workshop from *Elizabeth Holmes*.

Throughout this document, we will use `R` code to update to a new version of GMACS. The functions used are available in the [`gmr`] package, which is useful for working with GMACS. 
This package is available on the GMACS-project GitHub organisation. To get it installed you will need the [`devtools`] package. 


In the course of this document, we will also need various packages required to i) call AD Model Builder ([ADMB]) in `R` (to compile and build the GMACS executable). These packages are automatically loaded when loading the [`gmr`] package into R.

The following code proceeds the installation/update of the `[gmr]` package on your machine and load the various libraries.

```{r Install gmr and load packages}

rm(list=ls())     # Clean your R session

# Set the working directory as the directory of this document----
# setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

# check your directory
getwd()

# Install and load the packages----

# 1.Install devtools and gdata on your machine ----
if (!require("devtools")) {                        # install devtools
  install.packages("devtools")
}

if (!require("gdata")) {                        # needed to manipulate data
  install.packages("gdata")
}

# 2. Install / update gmr package ----
.Src <- "GMACS-project/gmr"
.Update <- 0                            # Get the latest version of gmr? (0: no; 1: install for the first time;
                                        # 2: update the package)
mylib <- "~/R/win-library/4.1"          # the library directory to remove the gmr package from 

#  remotes::install_github() will work to install gmr on your machine
if(.Update == 1) devtools::install_github(.Src)


# Updating to the latest version of gmr
if(.Update == 2){
  remove.packages("gmr", lib=mylib)
  devtools::install_github(.Src)
} 


# Load the gmr package
library(gmr)
```


`Codes` in this document are summarized in a clean and runnable `.R` script ([UpdateGMACS.R]) which you should use whenever you want to release a new version of GMACS. This R script will guide you through all the steps of the procedure making the GMACS upgrade process easier and more transparent.

# Set-up for using this workflow

To submit a new version of GMACS, you will need a [Github] account if you want to work with the GMACS-proect repositories and act as an active developer. 
You will also need to get installed R (or Rstudio).
Below are the links to install these programs:

-   Install [R]
-   Install [Rstudio]

If you want to link your R/Rstudio to your Github account so that you can push your changes from Rstudio directly to Github, please refer to the two workshops listed above. I will not cover this topic in this document. Noreover, in this document I will use [Github Desktop] to make the link between your updates on your local repositories and the one on your GitHub account.

# Let's get the *GMACS_Assessment_code* repository on your computer

I am going to show you a workflow to get the [GMACS_Assessment_code] repository on your computer. 
Here, you have two options:

1.  You want to contribute to the development of GMACS: you will need to [**fork**](#Fork-repo) the repository to be able to submit your changes to the organization owners;
2.  You only want to [**clone**](#Copy-repo) the repository of the organization and modify it for your own purpose (without contributing to the development of GMACS).

## Forking the [GMACS_Assessment_code] repository {#Fork-repo}

By ***forking*** this repository, you will be able to contribute to the organization's repository. 
If you just want to copy it and then modify it for your own purpose on you computer, 
please follow the steps described in the [Copy a GMACS-project repository](#Copy-repo) section.

1.  In [Github Desktop], click *File > Clone Repository*
2.  In the *"Clone a repository"* popup window, chose the URL tab and paste the 
GMACS_Assessment_code repository url.
3.  Check the selected folder in the local path and click `Clone`.

You are now ready to make some changes to GMACS, commit them and push up them to
your local repository. Asoon as you are ready to submit a new version of GMACS, you
will have to sumbit a Pull request on the GitHub organization.

## Clone the [GMACS_Assessment_code] repository {#Copy-repo}

Here I show you how to copy the [GMACS_Assessment_code] repository on your computer. 
By following these steps, you will not be able to contribute to the [GMACS-project] 
organisation but you will have the possibility to adapt this repository for your 
purpose. Below are the steps to follow:

1.  Get and copy the [url](https://github.com/GMACS-project/GMACS_Assessment_code) 
of the [GMACS_Assessment_code] repository;
2.  Go to you Github account. It should be github.com/**your-name**;
3.  In Github, click the `+` in top right and select `import repository`;
4.  Paste the url in ***the Your old repository's clone URL*** section and give 
a name for this new repo. You have now the [GMACS_Assessment_code] on your own 
Github (i.e., you have an url looking like *github.com/your-name/GMACS_Assessment_code* 
if you kept the same name as the one of the organization);
5.  You can now clone this repository to your computer following the same steps 
as those used to clone the [GMACS_Assessment_code] repository from the [organization] 
website.

# Let's work on a new version of GMACS

In the **[GMACS]** folder of the [GMACS_Assessment_code] repository, two subfolders 
should attract your attention when you want to implement a new version of GMACS 
(I'm talking about development when you make a change (no matter how small) to the GMACS code):

1.  The **[Dvpt_Version]** folder which contains the latest version of GMACS and where ***you should make your changes***. 
2.  The **[Latest_Version]** folder which holds the latest stable and tested 
version of GMACS (similar to the one in the [Dvpt_Version] folder).

These two subfolders contain all the hardware you need to run GMACS either in a 
"development way" or to get a "stable" stock assessment.

**If you want to make new changes to the GMACS code, you should work in the *[Dvpt_Version]* subfolder to keep the latest version "clean" till you run the comparison before submitting your changes.**

In the following, I give you some guidelines about how correctly update a GMACS 
version and make a pull request on Github. Obviously, this assumes that you have 
previously forked (and not cloned) the organization's repo on your computer.

Let's take the example of incorporating time-varying natural mortality into GMACS 
with a focus on snow crab. I will not go through the implementation of the code 
itself but simply give you the path to follow to:

i.  keep track of these changes,
ii. check that they do not impact the obtainment of a new executable (i.e., that the code can be compiled),
iii. analyze the impact of these changes on the assessment for a stock (i.e., compare with the last stable version of GMACS the results of the assessment using this version under development) and,
iv. release this new stable version of GMACS to the community.

*Reminder:* **You will be working in the [Dvpt_Version] subfolder.**

## Modifying the gmacsbase.tpl

In the [Dvpt_Version] subfolder, open the gmacsbase.tpl file in an editor and 
incorporate the new functions/variables to implement time-varying natural mortality. 
This operation does not require any "new entries" or variable declarations in the 
.ctl, .dat or .prj files so that the change you make will remain (first) at the 
gmacsbase.tpl file level.

## Check compilation and build the executable

Once you are done with the configuration of time-varying natural mortality in GMACS, 
you now need to check that you are still able to compile the model and build the
executable. To build the source files of GMACS into an executable, you will need 
to call ADMB. The only requirement is to provide GMACS with specific directories 
so it will be able locate ADMB and a C/C++ compiler.

To provide such directories, you can use the [ADpaths_Windows.txt] (OR ADpaths_MacOS on linux) 
file [GMACS_Versions] folder of the GMACS organizaltion. (You will need to modify 
the paths accordingly with your setups on your computer.) 

This text file holds the information of specific R variable names and the pathway to use them.
**Please do not modify the R variable names in this file otherwise you will have some troubles and will not be able to build the GMACS executable.** 

When build a new GMACS executable, you will be asked to provide a name for the new version 
of GMACS. Please check the name of the current GMACS version and provide one accordingly.

The following code compiles and build the GMACS executable.

```{r compile and build GMACS, include=TRUE}

# Define the name of the file containing the different pathways needed to build
# the GMACS executable 
ADMBpaths <- "ADpaths.txt"

# Run the GetGmacsExe function
.GetGmacsExe()

```

While the command is running, you will see on the console exactly what is going on.
If the compilation worked, you should have a new executable named `gmacs.exe` in 
the the [Dvpt_Version] directory. You are now ready to test this new version. 

## Modify the .dat, .ctl, .prj files

In our case, we only need to modify the control parameters for the Time varying 
natural mortality rates.
In the .ctl file, we have to change the *type* of model for natural mortality, 
the *phase of estimation*, the *standard deviation* of the deviations for the 
random walk, the number of  *step-changes*  (time block) for which we need to 
define the specific years position of the knots. Finally we specified the values 
for the initial parameters, their low bound, their high bound and their phase. 

## Run the Development version using a case study: Snow crab

You are now going to run the GMACS development version for a specific case study. 

First define the characteristics of this analysis:

```{r Run the development version}
# II- Run the development version----

# Define the working directories
# Set the working directories:
Dir_Dvpt_Vers <- file.path(getwd(), "Dvpt_Version")
Dir_Last_Vers <- file.path(getwd(), "Latest_Version")

# Stock of interest
Spc <-c(
  "SNOW_crab"
)

# Names of the GMACS version to consider
GMACS_version <- c(
  "Dvpt_Version"
)

# Define directories
VERSIONDIR <- Dir_Last_Vers

# Use Last Assessment for comparison?
# If yes, you must provide the names of the model for each species in the variable .ASSMOD_NAMES
# Those model folder must have to be hold in the folder Assessments
ASS <- FALSE

# Need to conpile the model?
# vector of length(.GMACS_version)
# 0: GMACS is not compiled. This assumes that an executable exists in the directory of the concerned version.
# 1: GMACS is compiles
COMPILE <- 0       # You already compile and build the executable

# Run GMACS
RUN_GMACS <- TRUE

# Use latest available data for the assessment?
LastAssDat <- TRUE

# Define the directories for ADMB
ADMBpaths <- "ADpaths.txt"

# Show Rterminal
VERBOSE <- TRUE

# Do comparison?
MAKE_Comp <- FALSE


```

Then, run the model:

```{r Run GMCS dvpt_version, include=TRUE}
res <- GMACS(
  Spc = .Spc,
  GMACS_version = GMACS_version,
  Dir = VERSIONDIR,
  ASS = ASS,
  compile = COMPILE,
  run = RUN_GMACS,
  LastAssDat = LastAssDat,
  ADMBpaths = ADMBpaths,
  make.comp = MAKE_Comp,
  verbose = VERBOSE
)
```

All the output files resulting from this run are stored in the **SNOW_M_time_varying** directory. 
You are now ready to compare the outputs of this GMACS development version with the ones
coming from the latest assessment.

## Compare the results of this new version with the results from the last assessment

In the [GMACS_Assessment_code] repository, the required input files and specific 
outputs files from the latest assessment for all stocks are stored in the 
folder [Assessments]. For each stocks, this folder contains the models that have 
been tested, presented and selected by the Crab Plan Team and the Scientific and 
Statistical Committee to realize the assessment of each crab stock.

This comparison can be done using the [Compare_Version_VS_Last_Assessment] 
Rmarkdown document. Here, we will compare the results of the stock assessment for snow crab between the latest assessment and the development version of GMACS. 
In the same way as previously, the comparison is done by calling the the `GMACS()` function. You should have been able to do the run and the comparison at the same time i.e., calling the `GMACS()` function only once but for the purpose of this document we split the steps. Obviously, we are not going again through the entire run here. You have first to modify the `.GMACS_version` and the `.VERSIONDIR` variables to consider the last assessment and turn off the `.RUN_GMACS` variable to avoid going again through the run. You will also need to specify the `.ASS` and `.ASSMOD_NAMES` variables that indicate you want to consider the last assessment in the comparison and the name of the model used for this assessment and of course, you will have to turn on the `.MAKE_Comp` variable to inticate that you want to make comparison between the two versions.

Changes the parameters for the comparison:

```{r, Compare the development version and the latest assessment}

# Names of the GMACS version to consider for run
.GMACS_version <- c(
  "Last_Assessment",
  "Dvpt_Version"
)

# Define directory
.VERSIONDIR <- c(
  paste0(dirname(getwd()), "/Assessments/"),
  paste0(getwd(), "/Dvpt_Version/")
)

# Need to conpile the model?
# vector of length(.GMACS_version)
# 0: GMACS is not compiled. This assumes that an executable exists in the directory of the concerned version.
# 1: GMACS is compiles
.COMPILE <- c(0,0)       # You already compile and build the executable

# Run GMACS
.RUN_GMACS <- FALSE

# Species
.Spc <- c(
          'SNOW_crab'
          )

# Use Last Assessment for comparison?
# If yes, you must provide the names of the model for each species in the variable .ASSMOD_NAMES
# Those model folder must have to be hold in the folder Assessments
.ASS <- TRUE

# names of the model for the last assessment - Only useful if comparison is made.
# if all stocks are considered they have to be ordered as follow:
# "AIGKC/EAG" / "AIGKC/WAG" / "BBRKC" / "SMBKC" / "SNOW"
.ASSMOD_NAMES <- c(
                   'model_21_g'
                   )

# Do comparison?
.MAKE_Comp <- TRUE

```

Call the GMACS() function:

```{r}
# Call the GMACS() function:

tables <- GMACS(Spc = .Spc, GMACS_version = .GMACS_version,
                Dir = .VERSIONDIR,
                compile = .COMPILE,
                ASS = .ASS,
                AssMod_names = .ASSMOD_NAMES,
                run = .RUN_GMACS,
                make.comp = .MAKE_Comp)

```



If you are satisfied with the results of the comparison between these two versions 
of GMACS, you are now ready to formalize and submit this new version to spread 
it to the community.

The first step before submitting your new version on GitHub is to update the 
[Latest_Version] folder with the new code of GMACS.


## Copy the latest version to the Latest_Version folder

Luckily and for the sake of efficiency and transparency, you don't have to do 
anything by hand. 

The `UpdateGMACS()` function allows you to:

  i) Copy and paste all the files you used for the GMACS development version to the [Latest_Version] folder
  ii) Compile this new release version in the [Latest_Version] folder and get everything ready to use it

```{r}
# Use the UpdateGMACS function to copy and paste the last files in the Latest_Version
# directory
UpdateGMACS()
```



## Push up changes to the organization's repository

## Re-run the assessments with the latest version (optional)

